Statistical Inference of Transport Mechanisms and Long Time Scale Behavior from Time Series of Solute Trajectories in Nanostructured Membranes

### 2020-08-11 Ben_Manuscripts/hdphmm/Draft.tex
Some additions to put things in context.

  Overall, this work provides a promising way to simultaneously identify transport 
  mechanisms in nanoporous materials and project complex diffusive behavior on
  long time scales. Our enhanced understanding of the diverse range of solute behavior
  allows us to hypothesize design changes to LLC monomers aimed towards controlling the
  rates of solute passage, thus improving the selective performance of LLC membranes.

@@

%  We have already advanced our understanding of LLC monomer design in our previous work
%  by building stochastic time series models based on the chemical intuition gained
%  from qualitative MD studies of solute transport in an H\textsubscript{II} phase LLC
%  membrane.~\cite{coscia_chemically_2019,coscia_capturing_2020} 
  Based on the chemical intuition gained from our previous qualitative MD studies of 
  solute transport in an H\textsubscript{II} phase LLC membrane, we have contributed our
  own stochastic modeling approaches with goals similar to both TEKMC and DBP.~\cite{coscia_chemically_2019,coscia_capturing_2020}
  We identified three classes of trapping behavior which result in subdiffusion out to the microsecond 
  time scale. In our first approach to stochastically model this behavior, we treated 
  the system's dynamics as a sequence of anti-correlated hops between power law distributed periods of 
  entrapment, a framework called subordinated fractional Brownian or L\'evy
  motion.~\cite{thiel_weak_2014,teuerle_modeling_2013} In a second approach, we treated 
  solute motion as a Markov state model~\cite{pande_everything_2010} with state-dependent
  dynamics, where we parameterized the state transition probabilities between each of
  eight discrete states, defined by the observed trapping mechanisms, as well as the 
  solute dynamics within each of these states. We showed how one could use realizations
  of any stochastic model like these in order to predict macroscopic flux and selectivity.

@@

  In this work we attempt to uncover similar physical meaning through a data-driven process. 
  We utilize the sticky hierarchical Dirichlet process autoregressive hidden Markov model 
  (HDP-AR-HMM), a nonparameteric Bayesian classification technique that allows the user to 
  automatically detect and infer the parameters of an unknown number of latent vector
  autoregressive (VAR) modes present in solute center-of-mass MD 
  trajectories.~\cite{fox_bayesian_2010} The sticky HDP-AR-HMM greatly extends the 
  basic framework of simpler but more limited finite hidden Markov modeling approaches.

@@

% LocalWords:  BJC micropollutants Desalination permeability nanofiltration LLC
% LocalWords:  Amphiphilic nanostructures Lyotropic amphiphilic lyotropic MSDDM
% LocalWords:  solutes selectivities solute nanoscopic timescales MSDs MSD HMMs
% LocalWords:  nonparameteric Parrinello Rahman barostat rescale transitioning
% LocalWords:  al's HMM Dirichlet HDP DP Wishart TODO SI al MATLAB teh gael pdf
% LocalWords:  nonparametric bayesian solute's reparameterized intracluster ca
% LocalWords:  delocalize Luzar Chandler histogrammed interpolator methanol's
% LocalWords:  nanostructure carboxylate VMD nonidealities hbonds outlier GAANN
% LocalWords:  autogregressive PRF ACS XSEDE ACI HDP-AR-HMM nanoporous flowback evy
% LocalWords:  timescale subdiffusive coscia subdiffusion calderon GROMACS der
% LocalWords:  berendsen gromacs spoel hess emailed plateauing hmm ne gelman AR
% LocalWords:  walt numpy GCL URE ACH IMMM subsubsection modT convective vmd Na
% LocalWords:  scannable subsubcaptions errorbars uncharacteristicness facelift
% LocalWords:  PSC RMACC hdphmm TOC nominalizations dischinger ramaseshan Baum
% LocalWords:  functionalized teuerle jasra markov bioinformatics hughes yoon
% LocalWords:  hines diarization MJLSs MJLS SLDSs SLDS hamada jk params MNIW BP
% LocalWords:  plateaued nclusters leftmargin trans diffusivity sorption alkane
% LocalWords:  ness substituents desalination fonseca couto welch singhal noe
% LocalWords:  shirtsgroup superdiffusive diffusivities jazani wang persson CPC
% LocalWords:  shortlag jupyter ester

>>>>

  This work provides a promising way to simultaneously identify transport 
  mechanisms in nanoporous materials and project complex diffusive behavior on
  long time scales. Our enhanced understanding of the diverse range of solute behavior
  allows us to hypothesize design changes to LLC monomers aimed towards controlling the
  rates of solute passage, thus improving the selective performance of LLC membranes.

@@

%  We have already advanced our understanding of LLC monomer design in our previous work
%  by building stochastic time series models based on the chemical intuition gained
%  from qualitative MD studies of solute transport in an H\textsubscript{II} phase LLC
%  membrane.~\cite{coscia_chemically_2019,coscia_capturing_2020} 
  Based on the chemical intuition gained from our previous qualitative MD studies of 
  solute transport in an H\textsubscript{II} phase LLC membrane, we have contributed our
  own stochastic modeling approaches with goals similar to both TEKMC and DBP.~\cite{coscia_chemically_2019,coscia_capturing_2020}
  In previous studies, we identified three classes of trapping behavior which result in subdiffusion out to the microsecond 
  time scale. In our first approach to stochastically model this behavior, we treated 
  the system's dynamics as a sequence of anti-correlated hops between power law distributed periods of 
  entrapment, a framework called subordinated fractional Brownian or L\'evy
  motion.~\cite{thiel_weak_2014,teuerle_modeling_2013} In a second approach, we treated 
  solute motion as a Markov state model~\cite{pande_everything_2010} with state-dependent
  dynamics, where we parameterized the state transition probabilities between each of
  eight discrete states, defined by the observed trapping mechanisms, as well as the 
  solute dynamics within each of these states. We showed how one could use realizations
  of any stochastic model like these in order to predict macroscopic flux and selectivity.

@@

  In this work we attempt to uncover similar physical meaning through
  a data-driven process in which we characterize the local behavior of
  the system through analysis of the dynamical characteristics o
  particle trajectories.  This differs from other other approaches
  such as TEKMC and DBP than cluster based on configurations visited
  to extend the trajectories. We utilize the sticky hierarchical
  Dirichlet process autoregressive hidden Markov model (HDP-AR-HMM), a
  nonparameteric Bayesian classification technique that allows the
  user to automatically detect and infer the parameters of an unknown
  number of latent vector autoregressive (VAR) modes present in solute
  center-of-mass MD trajectories.~\cite{fox_bayesian_2010} The sticky
  HDP-AR-HMM greatly extends the basic framework of simpler but more
  limited finite hidden Markov modeling approaches.

@@

% LocalWords:  BJC micropollutants Desalination permeability nanofiltration LLC
% LocalWords:  Amphiphilic nanostructures Lyotropic amphiphilic lyotropic MSDDM
% LocalWords:  solutes selectivities solute nanoscopic timescales MSDs MSD HMMs
% LocalWords:  nonparameteric Parrinello Rahman barostat rescale transitioning
% LocalWords:  al's HMM Dirichlet HDP DP Wishart TODO SI al MATLAB teh gael pdf
% LocalWords:  nonparametric bayesian solute's reparameterized intracluster ca
% LocalWords:  delocalize Luzar Chandler histogrammed interpolator methanol's
% LocalWords:  nanostructure carboxylate VMD nonidealities hbonds outlier GAANN
% LocalWords:  autogregressive PRF ACS XSEDE ACI HDP-AR-HMM nanoporous flowback evy
% LocalWords:  timescale subdiffusive coscia subdiffusion calderon GROMACS der
% LocalWords:  berendsen gromacs spoel hess emailed plateauing hmm ne gelman AR
% LocalWords:  walt numpy GCL URE ACH IMMM subsubsection modT convective vmd Na
% LocalWords:  scannable subsubcaptions errorbars uncharacteristicness facelift
% LocalWords:  PSC RMACC hdphmm TOC nominalizations dischinger ramaseshan Baum
% LocalWords:  functionalized teuerle jasra markov bioinformatics hughes yoon
% LocalWords:  hines diarization MJLSs MJLS SLDSs SLDS hamada jk params MNIW BP
% LocalWords:  plateaued nclusters leftmargin trans diffusivity sorption alkane
% LocalWords:  ness substituents desalination fonseca couto welch singhal noe
% LocalWords:  shirtsgroup superdiffusive diffusivities jazani wang persson CPC
% LocalWords:  shortlag jupyter ester wt TEKMC DBP Webb Antechamber
% LocalWords:  polarizable unequilibrated traj bicontinuous
